Simulating High-Dimensional Multivariate Data

### 2021-03-01 030-bigsimr-package.Rmd

This section describes a low-dimensional (2D) random vector simulation workflow via the `bigsimr` `R` package (https://github.com/SchisslerGroup/r-bigsimr).
The package `bigsimr` provides an interface to the native code written in Julia, registered as the `Bigsimr` Julia package (https://github.com/adknudson/Bigsimr.jl).
In addition to the native Julia `Bigsimr` package and `R` interface `bigsimr`, we also provide a python interface `bigsimr` (https://github.com/SchisslerGroup/python-bigsimr/) that interfaces with the Julia `Bigsimr` package. 
The Julia implementation provides a high-performance implementation of our proposed random vector generation algorithm and associated functions (see Section \@ref(algorithms)).
When designing `bigsimr`, we aimed to conveniently provide parallelized computation, through multi-core and GPU acceleration, while allowing advanced users to customize.

The subsections below describe the basic use of `bigsimr`, by stepping through an example workflow using the data set `airquality` that contains daily air quality measurements in New York, May to September 1973 [@Chambers1983].
This workflow proceeds from computing environment setup, data wrangling, estimation, simulation configuration, random vector generation, and, finally, result visualization.
The code chunk below prepares the computing environment:

<!-- 
`bigsimr` is an R package for simulating high-dimensional multivariate data with arbitrary marginal distributions. The efficiency behind generating multivariate samples comes from the ability to utilize a GPU during the generation of multivariate normal samples and the subsequent transformation to uniform samples (see section \@ref(rand-vec-gen)). A GPU is not necessary in order to use `bigsimr`, and the speed benefits generally only come when simulating very high dimensional data ($d > 10000$).

The next computational step after obtaining the correlated uniform marginals is the inverse transform into the target marginal distributions. Since this consists of $d$ independent transformations, we utilize parallelization to achieve higher throughput. As with many parallel algorithms, there is overhead associated with forking the task to utilize multiple cores, however cost is negligible compared to the total run-time.
-->

@@

We illustrate the use of `bigsimr` motivated by the New York air quality data set (`airquality`) included in the R `datasets` package.
First, we load the `bigsimr` library and a few other convenient data science packages, including the syntactically-elegant `tidyverse` suite of `R` packages.

@@

Next, we specify the marginal distributions and correlation coefficient (both type and magnitude). Here the analyst is free to be creative. For this example, we avoid goodness-of-fit considerations to determine the marginal distributions. But it seems sensible without domain knowledge to estimate these quantities from the data and `bigsimr` contains fast functions designed for this task.

@@

As mentioned, the user must decide how to describe correlation, based on the particulars of the problem. For non-normal data and for improved simulation accuracy/scalability in our scheme, we advocate the use of Spearman's $\rho$ correlation matrix $R_S$ and Kendall's $\tau$ correlation matrix $R_K$. We also support Pearson correlation coefficient matching, while cautioning the user to check the performance for their parametric multivariate model (see [Monte Carlo evaluations](#simulations) below for evaluation strategies and guidance). These estimation methods are classical approaches, not designed for high-dimensional correlation estimation (see the [Conclusion and Discussion]({#discussion) sections for more on this).

@@

```{r ch030-cor-bounds, echo=TRUE}
bounds <- bs$cor_bounds(margins)
bounds$lower
```

@@

## Simulating random vectors
Finally, we arrive at the main function of `bigsimr`, `rvec`. We now simulate $B=10,000$ random vectors from the assumed joint distribution of Ozone levels and Temp.

>>>>

This section describes a low-dimensional (2D) random vector simulation workflow via the `bigsimr` `R` package (https://github.com/SchisslerGroup/r-bigsimr). The package `bigsimr` provides an interface to the native code written in Julia, registered as the `Bigsimr` Julia package (https://github.com/adknudson/Bigsimr.jl). In addition to the native Julia `Bigsimr` package and `R` interface `bigsimr`, we also provide a python interface `bigsimr` (https://github.com/SchisslerGroup/python-bigsimr/) that interfaces with the Julia `Bigsimr` package. The Julia package provides a high-performance implementation of our proposed random vector generation algorithm and associated functions (see Section \@ref(algorithms)).

The subsections below describe the basic use of `bigsimr` by stepping through an example workflow using the data set `airquality` that contains daily air quality measurements in New York, May to September 1973 [@Chambers1983]. This workflow proceeds from setting up the computing environment, to data wrangling, estimation, simulation configuration, random vector generation, and, finally, result visualization.

@@

We illustrate the use of `bigsimr` motivated by the New York air quality data set (`airquality`) included in the R `datasets` package. First, we load the `bigsimr` library and a few other convenient data science packages, including the syntactically-elegant `tidyverse` suite of `R` packages. The code chunk below prepares the computing environment:

```{r ch030-computeEnvironmentSetup, eval=FALSE, echo=TRUE}
library("tidyverse")
library("bigsimr")
# Activate multithreading in Julia
Sys.setenv(JULIA_NUM_THREADS = parallel::detectCores())
# Load the Bigsimr and Distributions Julia packages
bs <- bigsimr_setup()
dist <- distributions_setup()
```

@@

Next, we specify the marginal distributions and correlation coefficient (both type and magnitude). Here the analyst is free to be creative. For this example, we avoid goodness-of-fit considerations to determine the marginal distributions. But it seems sensible without domain knowledge to estimate these quantities from the data, and `bigsimr` contains fast functions designed for this task.

@@

As mentioned, the user must decide how to describe correlation based on the particulars of the problem. For non-normal data and for improved simulation accuracy/scalability in our scheme, we advocate the use of Spearman's $\rho$ correlation matrix $R_S$ and Kendall's $\tau$ correlation matrix $R_K$. We also support Pearson correlation coefficient matching, while cautioning the user to check the performance for their parametric multivariate model (see [Monte Carlo evaluations](#simulations) below for evaluation strategies and guidance). These estimation methods are classical approaches, not designed for high-dimensional correlation estimation (see the [Conclusion and Discussion]({#discussion) sections for more on this).

@@

```{r ch030-cor-bounds, echo=TRUE}
bounds <- bs$cor_bounds(margins)
bounds$lower
bounds$upper
```

@@

## Simulating random vectors

Finally, we arrive at the main function of `bigsimr`: `rvec`. We now simulate $B=10,000$ random vectors from the assumed joint distribution of Ozone levels and Temp.
